oREV: an Item Response Theory based open receptive vocabulary task for 3 to 8-year-old children

### 2022-08-13 paper/manuscript.Rmd
fix minor edits

Individual differences in language abilities are early emerging, stable across development, and predictive of a wide range of psychological outcome variables including cognitive abilities, academic achievement, or mental health [@bornstein2018stability; @schoon2010children; @morgan201524; @marchman2008speed; @walker1994prediction]. From a methodological perspective, high-quality, easy-access measures of language abilities are therefore central to both basic and applied research on individual differences in language abilities. Ideally, such measures should also be comparable across languages in order to study which developmental processes are language-specific and which are shared more widely. Developing such measures is very time and resource intensive and, as a consequence, few exist. In this paper, we describe the construction of a new receptive vocabulary task for German-speaking children. Its psychometric grounding in Item Response Theory makes the measure robust and efficient. Its web-based design and implementation make the measure easy to adapt and administer in different settings (in-person or remote) and thereby facilitate the scaling of data collection.

@@

A range of measures exist to assess vocabulary skills in children. For very young children (up to 3 years), a very popular instrument is the MacArthur–Bates Communicative Development Inventories (CDIs) [@fenson2007macarthur]. Parents are provided with a list of words and are asked to check those the child understands and/or produces. The CDI exists in different forms [e.g., @makransky2016item; @mayor2019short], including and online version [@demayo2021web], and has been adapted to many different languages [see @frank2021variability]. Thanks to concentrated collaborative efforts, data from thousands of children learning dozens of of languages has been pooled in centralized repositories [@frank2017wordbank; @jorgensen2010clex]. As such the CDI provides positive example for a high-quality, easy-access measure that is heavily used in both basic and applied research.

@@

The development of so-called Cross-linguistic Lexical Tasks (ClTs; @haman2015designing) constitutes a promising framework that might help to overcome these issues. CLTs are simple picture-choice and picture-naming tasks aimed at assessing comprehension and production of nouns and verbs. In a collaborative effort involving more than 25 institutions, versions for dozens of different languages have been developed following the same guiding principles [@haman2015designing; @haman2017noun; @armon2015assessing]. In addition to cross-linguistic studies with mono-lingual children, this procedure makes CLTs ideally suited to assess multi-lingual children. The tasks and the materials are not commercially licensed and can thus be freely used for research purposes.

@@

The initial item pool consisted of 32 items taken from the German CLT [@haman2015designing; @haman2017noun] and 20 new items. The addition of new items was necessary due to ceiling effects for monolingual 5-year-olds in the previous version. New items were generated in line with the construction of the original CLT in a stepwise process. Each item consists of a target word and three distractors. To select target words, we first compiled a list of age-of-acquisition ratings for 3,928 German words from various sources [@luniewska2019age; @schroder2012german; @birchenough2017rated]. From this list, we selected 20 words based on the following criteria: words should refer to concepts that could easily and unambiguously be depicted in a drawing, age-of-acquisition ratings should be spread equally between six and ten years of age. We also computed complexity indices for each word [see @haman2017noun]. This metric, however, did not reflect a dimension that was relevant for item selection.

@@

Individual differences in language abilities in childhood are an important predictor of later life outcomes. Yet, high-quality, easy access measures are rare, especially for preschool-aged children. Here we reported the construction of a new receptive vocabulary task for German-speaking children between 3 and 8 years-of-age. Building on earlier work [@haman2017noun], we first generated a larger initial pool with 52 items. Next, we implemented the picture-selection task as a web application and collected data from more than 500 children online. We used IRT models and an automated item selection algorithm to select a minimal set of high quality items. The so-constructed task has 20 items and correlates with the full task at a rate of .97. Its browser-based implementation makes the task highly portable and facilitates large scale data collection. The construction and item selection process we described here makes it easy to add additional items or adapt the task to different languages while retaining a high psychometric quality of the end product. The task is freely accessible to all interested researchers.

The task fills an important gap in the methods repertoire of developmental researchers studying language development in the preschool years. Existing measures show ceiling effects, come with high licensing costs and/or are not available in an electronic format. Our task captures variation between children up until 8 years of age, is free to use and can be run on any modern web browser. However, the newly constructed task with 20 items is still relatively easy, that is, most 7 year-old children will solve the majority of items (`r perf_sub_task%>%filter(age_group == 7)%>%pull(mean)*100`% correct responses in the present sample). As a consequence, it does not distinguish well between children with very strong vocabulary skills. Future extensions of the task could thus focus on adding more difficult items. Figure \@ref(fig:fig2)B [see also @brysbaert2017test] shows that target word age-of-acquisition ratings are a fairly good predictor of item difficulty and could be used as a basis to generate new items. Extensions should focus on target words with rated age-of-acquisition above 10.

>>>>

Individual differences in language abilities are early emerging, stable across development, and predictive of a wide range of psychological outcome variables including cognitive abilities, academic achievement, or mental health [@bornstein2018stability; @schoon2010children; @morgan201524; @marchman2008speed; @walker1994prediction]. From a methodological perspective, high-quality, easy-access measures of language abilities are therefore central to both basic and applied research on individual differences in language abilities. Ideally, such measures should also be comparable across languages in order to study which developmental processes are language-specific and which are shared more widely. Developing such measures is very time and resource intensive and, as a consequence, few exist. In this paper, we describe the construction of a new receptive vocabulary task for German-speaking children. Its psychometric grounding in Item Response Theory makes the measure robust and efficient. Its web-based design and implementation makes the measure easy to adapt and administer in different settings (in-person or remote) and thereby facilitates the scaling of data collection.

@@

A range of measures exist to assess vocabulary skills in children. For very young children (up to 3 years), a very popular instrument is the MacArthur–Bates Communicative Development Inventories (CDIs) [@fenson2007macarthur]. Parents are provided with a list of words and are asked to check those the child understands and/or produces. The CDI exists in different forms [e.g., @makransky2016item; @mayor2019short], including an online version [@demayo2021web], and has been adapted to many different languages [see @frank2021variability]. Thanks to concentrated collaborative efforts, data from thousands of children learning dozens of languages has been pooled in centralized repositories [@frank2017wordbank; @jorgensen2010clex]. As such the CDI provides positive example for a high-quality, easy-access measure that is heavily used in both basic and applied research.

@@

The development of so-called Cross-linguistic Lexical Tasks (CLTs; @haman2015designing) constitutes a promising framework that might help to overcome these issues. CLTs are simple picture-choice and picture-naming tasks aimed at assessing comprehension and production of nouns and verbs. In a collaborative effort involving more than 25 institutions, versions for dozens of different languages have been developed following the same guiding principles [@haman2015designing; @haman2017noun; @armon2015assessing]. In addition to cross-linguistic studies with mono-lingual children, this procedure makes CLTs ideally suited to assess multi-lingual children. The tasks and the materials are not commercially licensed and can thus be freely used for research purposes.

@@

The initial item pool consisted of 32 items taken with permission from the German CLT [@haman2015designing; @haman2017noun] and 20 new items. The addition of new items was necessary due to ceiling effects for monolingual 5-year-olds in the previous version. New items were generated in line with the construction of the original CLT in a stepwise process. Each item consists of a target word and three distractors. To select target words, we first compiled a list of age-of-acquisition ratings for 3,928 German words from various sources [@luniewska2019age; @schroder2012german; @birchenough2017rated]. From this list, we selected 20 words based on the following criteria: words should refer to concepts that could easily and unambiguously be depicted in a drawing, age-of-acquisition ratings should be spread equally between six and ten years of age. We also computed complexity indices for each word [see @haman2017noun]. This metric, however, did not reflect a dimension that was relevant for item selection.

@@

Individual differences in language abilities in childhood are an important predictor of later life outcomes. Yet, high-quality, easy access measures are rare, especially for pre- and primary-school-aged children. Here we reported the construction of a new receptive vocabulary task for German-speaking children between 3 and 8 years-of-age. Building on earlier work [@haman2017noun], we first generated a larger initial pool with 52 items. Next, we implemented the picture-selection task as a web application and collected data from over 500 children online. We used IRT models and an automated item selection algorithm to select a minimal set of high quality items. The so-constructed task has 20 items and correlates with the full task at a rate of .97. Its browser-based implementation makes the task highly portable and facilitates large scale data collection. The construction and item selection process we described here makes it easy to add additional items or adapt the task to different languages while retaining a high psychometric quality of the end product. The task is freely accessible to all interested researchers.

The task fills an important gap in the methods repertoire of developmental researchers studying monolingual and bilingual language development in early childhood. Existing measures show ceiling effects, come with high licensing costs and/or are not available in an electronic format. Our task captures variation between children up until 8 years of age, is free to use and can be run on any modern web browser. However, the newly constructed task with 20 items is still relatively easy, that is, most 7 year-old children will solve the majority of items (`r perf_sub_task%>%filter(age_group == 7)%>%pull(mean)*100`% correct responses in the present sample). As a consequence, it does not distinguish well between children with very strong vocabulary skills. Future extensions of the task could thus focus on adding more difficult items. Figure \@ref(fig:fig2)B [see also @brysbaert2017test] shows that target word age-of-acquisition ratings are a fairly good predictor of item difficulty and could be used as a basis to generate new items. Extensions should focus on target words with rated age-of-acquisition above 10. Further extension could other parts of speech, such as verbs and adjectives.

### 2022-08-13 paper/manuscript.tex
fix minor edits

Individual differences in language abilities are early emerging, stable across development, and predictive of a wide range of psychological outcome variables including cognitive abilities, academic achievement, or mental health (Bornstein, Hahn, Putnick, \& Pearson, 2018; Marchman \& Fernald, 2008; Morgan, Farkas, Hillemeier, Hammer, \& Maczuga, 2015; Schoon, Parsons, Rush, \& Law, 2010; Walker, Greenwood, Hart, \& Carta, 1994). From a methodological perspective, high-quality, easy-access measures of language abilities are therefore central to both basic and applied research on individual differences in language abilities. Ideally, such measures should also be comparable across languages in order to study which developmental processes are language-specific and which are shared more widely. Developing such measures is very time and resource intensive and, as a consequence, few exist. In this paper, we describe the construction of a new receptive vocabulary task for German-speaking children. Its psychometric grounding in Item Response Theory makes the measure robust and efficient. Its web-based design and implementation make the measure easy to adapt and administer in different settings (in-person or remote) and thereby facilitate the scaling of data collection.

@@

A range of measures exist to assess vocabulary skills in children. For very young children (up to 3 years), a very popular instrument is the MacArthur--Bates Communicative Development Inventories (CDIs) (Fenson et al., 2007). Parents are provided with a list of words and are asked to check those the child understands and/or produces. The CDI exists in different forms (e.g., Makransky, Dale, Havmose, \& Bleses, 2016; Mayor \& Mani, 2019), including and online version (DeMayo et al., 2021), and has been adapted to many different languages (see Frank, Braginsky, Yurovsky, \& Marchman, 2021). Thanks to concentrated collaborative efforts, data from thousands of children learning dozens of of languages has been pooled in centralized repositories (Frank, Braginsky, Yurovsky, \& Marchman, 2017; Jørgensen, Dale, Bleses, \& Fenson, 2010). As such the CDI provides positive example for a high-quality, easy-access measure that is heavily used in both basic and applied research.

@@

The development of so-called Cross-linguistic Lexical Tasks (ClTs; Haman, Łuniewska, and Pomiechowska (2015)) constitutes a promising framework that might help to overcome these issues. CLTs are simple picture-choice and picture-naming tasks aimed at assessing comprehension and production of nouns and verbs. In a collaborative effort involving more than 25 institutions, versions for dozens of different languages have been developed following the same guiding principles (Armon-Lotem, Jong, \& Meir, 2015; Haman et al., 2017, 2015). In addition to cross-linguistic studies with mono-lingual children, this procedure makes CLTs ideally suited to assess multi-lingual children. The tasks and the materials are not commercially licensed and can thus be freely used for research purposes.

@@

The initial item pool consisted of 32 items taken from the German CLT (Haman et al., 2017, 2015) and 20 new items. The addition of new items was necessary due to ceiling effects for monolingual 5-year-olds in the previous version. New items were generated in line with the construction of the original CLT in a stepwise process. Each item consists of a target word and three distractors. To select target words, we first compiled a list of age-of-acquisition ratings for 3,928 German words from various sources (Birchenough, Davies, \& Connelly, 2017; Łuniewska et al., 2019; Schröder, Gemballa, Ruppin, \& Wartenburger, 2012). From this list, we selected 20 words based on the following criteria: words should refer to concepts that could easily and unambiguously be depicted in a drawing, age-of-acquisition ratings should be spread equally between six and ten years of age. We also computed complexity indices for each word (see Haman et al., 2017). This metric, however, did not reflect a dimension that was relevant for item selection.

@@

Individual differences in language abilities in childhood are an important predictor of later life outcomes. Yet, high-quality, easy access measures are rare, especially for preschool-aged children. Here we reported the construction of a new receptive vocabulary task for German-speaking children between 3 and 8 years-of-age. Building on earlier work (Haman et al., 2017), we first generated a larger initial pool with 52 items. Next, we implemented the picture-selection task as a web application and collected data from more than 500 children online. We used IRT models and an automated item selection algorithm to select a minimal set of high quality items. The so-constructed task has 20 items and correlates with the full task at a rate of .97. Its browser-based implementation makes the task highly portable and facilitates large scale data collection. The construction and item selection process we described here makes it easy to add additional items or adapt the task to different languages while retaining a high psychometric quality of the end product. The task is freely accessible to all interested researchers.

The task fills an important gap in the methods repertoire of developmental researchers studying language development in the preschool years. Existing measures show ceiling effects, come with high licensing costs and/or are not available in an electronic format. Our task captures variation between children up until 8 years of age, is free to use and can be run on any modern web browser. However, the newly constructed task with 20 items is still relatively easy, that is, most 7 year-old children will solve the majority of items (87\% correct responses in the present sample). As a consequence, it does not distinguish well between children with very strong vocabulary skills. Future extensions of the task could thus focus on adding more difficult items. Figure \ref{fig:fig2}B (see also Brysbaert \& Biemiller, 2017) shows that target word age-of-acquisition ratings are a fairly good predictor of item difficulty and could be used as a basis to generate new items. Extensions should focus on target words with rated age-of-acquisition above 10.

>>>>

Individual differences in language abilities are early emerging, stable across development, and predictive of a wide range of psychological outcome variables including cognitive abilities, academic achievement, or mental health (Bornstein, Hahn, Putnick, \& Pearson, 2018; Marchman \& Fernald, 2008; Morgan, Farkas, Hillemeier, Hammer, \& Maczuga, 2015; Schoon, Parsons, Rush, \& Law, 2010; Walker, Greenwood, Hart, \& Carta, 1994). From a methodological perspective, high-quality, easy-access measures of language abilities are therefore central to both basic and applied research on individual differences in language abilities. Ideally, such measures should also be comparable across languages in order to study which developmental processes are language-specific and which are shared more widely. Developing such measures is very time and resource intensive and, as a consequence, few exist. In this paper, we describe the construction of a new receptive vocabulary task for German-speaking children. Its psychometric grounding in Item Response Theory makes the measure robust and efficient. Its web-based design and implementation makes the measure easy to adapt and administer in different settings (in-person or remote) and thereby facilitates the scaling of data collection.

@@

A range of measures exist to assess vocabulary skills in children. For very young children (up to 3 years), a very popular instrument is the MacArthur--Bates Communicative Development Inventories (CDIs) (Fenson et al., 2007). Parents are provided with a list of words and are asked to check those the child understands and/or produces. The CDI exists in different forms (e.g., Makransky, Dale, Havmose, \& Bleses, 2016; Mayor \& Mani, 2019), including an online version (DeMayo et al., 2021), and has been adapted to many different languages (see Frank, Braginsky, Yurovsky, \& Marchman, 2021). Thanks to concentrated collaborative efforts, data from thousands of children learning dozens of languages has been pooled in centralized repositories (Frank, Braginsky, Yurovsky, \& Marchman, 2017; Jørgensen, Dale, Bleses, \& Fenson, 2010). As such the CDI provides positive example for a high-quality, easy-access measure that is heavily used in both basic and applied research.

@@

The development of so-called Cross-linguistic Lexical Tasks (CLTs; Haman, Łuniewska, and Pomiechowska (2015)) constitutes a promising framework that might help to overcome these issues. CLTs are simple picture-choice and picture-naming tasks aimed at assessing comprehension and production of nouns and verbs. In a collaborative effort involving more than 25 institutions, versions for dozens of different languages have been developed following the same guiding principles (Armon-Lotem, Jong, \& Meir, 2015; Haman et al., 2017, 2015). In addition to cross-linguistic studies with mono-lingual children, this procedure makes CLTs ideally suited to assess multi-lingual children. The tasks and the materials are not commercially licensed and can thus be freely used for research purposes.

@@

The initial item pool consisted of 32 items taken with permission from the German CLT (Haman et al., 2017, 2015) and 20 new items. The addition of new items was necessary due to ceiling effects for monolingual 5-year-olds in the previous version. New items were generated in line with the construction of the original CLT in a stepwise process. Each item consists of a target word and three distractors. To select target words, we first compiled a list of age-of-acquisition ratings for 3,928 German words from various sources (Birchenough, Davies, \& Connelly, 2017; Łuniewska et al., 2019; Schröder, Gemballa, Ruppin, \& Wartenburger, 2012). From this list, we selected 20 words based on the following criteria: words should refer to concepts that could easily and unambiguously be depicted in a drawing, age-of-acquisition ratings should be spread equally between six and ten years of age. We also computed complexity indices for each word (see Haman et al., 2017). This metric, however, did not reflect a dimension that was relevant for item selection.

@@

Individual differences in language abilities in childhood are an important predictor of later life outcomes. Yet, high-quality, easy access measures are rare, especially for pre- and primary-school-aged children. Here we reported the construction of a new receptive vocabulary task for German-speaking children between 3 and 8 years-of-age. Building on earlier work (Haman et al., 2017), we first generated a larger initial pool with 52 items. Next, we implemented the picture-selection task as a web application and collected data from over 500 children online. We used IRT models and an automated item selection algorithm to select a minimal set of high quality items. The so-constructed task has 20 items and correlates with the full task at a rate of .97. Its browser-based implementation makes the task highly portable and facilitates large scale data collection. The construction and item selection process we described here makes it easy to add additional items or adapt the task to different languages while retaining a high psychometric quality of the end product. The task is freely accessible to all interested researchers.

The task fills an important gap in the methods repertoire of developmental researchers studying monolingual and bilingual language development in early childhood. Existing measures show ceiling effects, come with high licensing costs and/or are not available in an electronic format. Our task captures variation between children up until 8 years of age, is free to use and can be run on any modern web browser. However, the newly constructed task with 20 items is still relatively easy, that is, most 7 year-old children will solve the majority of items (87\% correct responses in the present sample). As a consequence, it does not distinguish well between children with very strong vocabulary skills. Future extensions of the task could thus focus on adding more difficult items. Figure \ref{fig:fig2}B (see also Brysbaert \& Biemiller, 2017) shows that target word age-of-acquisition ratings are a fairly good predictor of item difficulty and could be used as a basis to generate new items. Extensions should focus on target words with rated age-of-acquisition above 10. Further extension could other parts of speech, such as verbs and adjectives.
